SU($n$) ferromagnets near the Heisenberg point

### 2020-06-23 collective_sun/sun_perturbation.tex
rearranged discussion of truncated shell simulations added an intro for the appendix on multiplying multi-body operators

% todo: solve collective model; Casmir operators?

@@

In this work, we generalize some of the techniques used to harness
SU(2)-symmetric spin interactions for metrological advantage to the
case of SU($n$).  In doing so, we develop a low-energy effective
theory of gapped SU($n$) ferromagnets with weakly broken SU($n$)
symmetry (e.g.~due to external fields or additional interaction
terms).  The restriction of our theory to ferromagnetic interactions
can be relaxed to the assumption of
\begin{enumerate*}
\item an initial state in the manifold of permutationally-symmetric
  states, a multilevel generalization of the Dicke manifold for
  two-level spins, together with the assumption that
\item this manifold is gapped away from states that break
  permutational symmetry.
\end{enumerate*}
In addition to working out the effective Hamiltonian induced by weak
SU($n$) symmetry breaking through second order in perturbation theory,
we develop a diagrammatic method to systematically compute matrix
elements of arbitrary spin operators within manifolds of weakly broken
permutational symmetry, thereby enabling the numerical simulation of
dynamics induced by non-perturbative modifications to the SU($n$)
Heisenberg model.  While these simulations generally accumulate
uncontrolled errors from the truncation of Hilbert space, we identify
criteria under which numerical results correctly capture the
qualitative dynamical behavior of a system.

@@

If the coefficients $h_{ij}$ of the interaction Hamiltonian $H_0$ are
all negative, then the ground-state manifold $\M_0$ of $H_0$ consists
of spin-polarized, {\it permutationally symmetric} (PS) states that
are simultaneous $+1$ eigenstates of all permutation operators
$\Pi_{ij}$.  We will assume that all $h_{ij}<0$ throughout these
notes, but note that this restriction can be relaxed to the assumption
that the initial state of all spins is a permutationally-symmetric
state (e.g.~a spin-polarized state).  We will also assume that the PS
manifold $\M_0$ is gapped away from all orthogonal states by an
interaction energy $\Delta_{\t{gap}}>0$.  Power-law couplings of the
form $h_{ij}\sim-1/\abs{i-j}^\alpha$ on a $D$-dimensional lattice, for
example, always yield a non-vanishing spectral gap $\Delta_{\t{gap}}$
for long-range interactions with $\alpha\le D$ (see Appendix
\ref{sec:spectral_gap}).

@@

If the addition $\O\p{\m w,X}$ to the SU($n$)-symmetric Hamiltonian
$H_0$ in \eqref{eq:H_0} is sufficiently small, namely with operator
norm $\norm{\O\p{\m w,X}}$ less than half the spectral gap
$\Delta_{\t{gap}}$ of $H_0$, $\norm{\O\p{\m w,X}}<\Delta_{\t{gap}}/2$,
then we can treat the effect of $\O\p{\m w,X}$ on the ground-state PS
manifold $\M_0$ perturbatively.  The effective Hamiltonians
$H_{\t{eff}}^{(1)}$ and $H_{\t{eff}}^{(2)}$ induced by $\O\p{\m w,X}$
on $\M_0$ at first and second order in perturbation theory
are\cite{bravyi2011schrieffer}
\begin{align}
  H_{\t{eff}}^{(1)} = \P_0 \O\p{\m w,X} \P_0,
  &&
  H_{\t{eff}}^{(2)} = - \sum_{\Delta\ne0}
  \f1\Delta \P_0 \O\p{\m w,X} \P_\Delta \O\p{\m w,X} \P_0,
  \label{eq:H_eff}
\end{align}
where $\P_\Delta$ is a projector onto the eigenspace $\E_\Delta$ of
the interaction Hamiltonian $H_0$ with interaction energy $\Delta$
above that of PS manifold $\M_0$.  In particular, $\P_0$ is a
projector onto $\M_0$ itself, and $\E_0=\M_0$.  The first order
effective Hamiltonian $H_{\t{eff}}^{(1)}$ simply projects
$\O\p{\m w,X}$ onto the PS manifold $\M_0$, and takes the form
\begin{align}
  H_{\t{eff}}^{(1)} \EQPS \mean w \, \col X,
  \label{eq:H_eff_1}
\end{align}
where $\EQPS$ denotes equality under a restriction to $\M_0$; the
coefficient $\mean w$ is the average of all $w_k$; and $\col X$ is a
collective version of $X$:
\begin{align}
  \mean w \equiv \f1{\abs{\D_N\p{M}}}
  \sum_{k\in\D_N\p{M}} w_k,
  &&
  \col X \equiv \sum_{k\in\D_N\p{M}} X_k,
  \label{eq:mean_col}
\end{align}
with $\abs{\D_N\p{M}}=\prod_{j=0}^{M-1}\p{N-j}$.  In the case of a
magnetic field $s_\z$ or Ising interactions $s_\z\otimes s_\z$, for
example,
\begin{align}
  \col{s_\z} = \sum_i s_\z^{(i)} = S_\z,
  &&
  \col{s_\z\otimes s_\z}
  = \sum_{i\ne j} s_\z^{(i)} s_\z^{(j)}
  = S_\z^2 - \f14 N.
\end{align}
The second order effective Hamiltonian $H_{\t{eff}}^{(2)}$ in
\eqref{eq:H_eff} takes more work to simplify due to the presence of a
projector $\P_\Delta$ onto the manifold $\E_\Delta$ of states with
excitation energy $\Delta$.  This projector essentially picks off the
part of $\O\p{\m w,X}$ that is strictly off-diagonal with respect to
the ground- and excited-state manifolds $\E_0$ and $\E_\Delta$.  We
therefore need to decompose $\O\p{\m w,X}$ into components that
generate states of definite excitation energy when acting on PS states
$\ket\psi\in\M_0$.  The SU($n$) symmetry of $H_0$ enables such a
decomposition to take the form
\begin{align}
  H_0 \O\p{\m w,X} \ket\psi
  = \sum_\Delta \p{E_0+\Delta} \O\p{\m w_\Delta,X} \ket\psi,
\end{align}
where, thinking of the tensor $\m w$ as a $\abs{\D_N\p{M}}$-component
vector, the tensor $\m w_\Delta$ can be found by diagonalizing a
matrix $\hat{\m h}$ of dimensions
$\abs{\D_N\p{M}}\times\abs{\D_N\p{M}}\sim N^M\times N^M$, and
projecting $\m w$ onto the eigenspace of $\hat{\m h}$ with eigenvalue
$\Delta$ (see Appendix \ref{sec:eigenstates}).  The matrix
$\hat{\m h}$ is determined entirely by the couplings couplings
$h_{ij}$ of $H_0$.  While the computational cost of diagonalizing
$\hat{\m h}$ is nominally $O\p{N^{3M}}$, spatial symmetries of $H_0$
and $\O\p{\m w,X}$ such as translational invariance or isotropy can
drastically reduce the complexity of this eigenvalue problem, as we
discuss in Appendix \ref{sec:symmetries}.

@@

In order to find the effective Hamiltonian $H_{\t{eff}}^{(2)}$ induced
by the $M$-body operator $\O\p{\m w,X}$ on the PS manifold $\M_0$ at
second order in perturbation theory, we constructed states with
definite excitation energy $\Delta>0$ with respect to $H_0$.  Rather
than treating these excited states as ``intermediate'' or ``virtual''
in the derivation of an effective ground-state Hamiltonian, we can
perform numerical simulations in the joint Hilbert space of both
ground and excited states.  When the perturbation $\O\p{\m w,X}$ is
small, namely with operator norm
$\norm{\O\p{\m w,X}}<\Delta_{\t{gap}}/2$, such simulations are
equivalent (up to vanishing perturbative corrections) to a
higher-order perturbation theory.  Moreover, the first-class treatment
of excited states in finite systems enables simulating the dynamics
induced by {\it non-perturbative} additions $\O\p{\m w,X}$ to the base
Hamiltonian $H_0$.

In order to simulate dynamics within excited-state manifolds, we first
need to classify the set of excited states, and construct a suitable
basis.  In the case of SU(2), a common starting point is to identify
states with a definite (non-negative) spin length
$S=\set{N/2,N/2-1,\cdots}$, and a spin projection
$m_\z\in\set{S,S-1,\cdots,-S}$ onto a quantization axis.  To
generalize this classification to SU($n$), we denote the manifold of
states with definite fixed spin length $S$ by $\M_{N/2-S}$.  The PS
manifold of Dicke states with $S=N/2$ is then $\M_0$, the manifold of
single-magnon spin-wave states with $S=N/2-1$ is $\M_1$, and more
generally the manifold of $r$-magnon states with $S=N/2-r$ is $\M_r$.
Rather than the spin projection $m_\z$, we then identify states by
labels $m=\p{m_\up,m_\dn}\in\A_2\p{N}$ in which $m_\mu$ denotes the
occupation number of the single-spin state $\mu$, with $m_\up+m_\dn=N$
and $m_\up-m_\dn=m_\z$.

% \begin{figure}
%   \centering
%   \includegraphics{structure/mat_L101_a2.pdf}
%   \caption{Visualizations of dimension-$r$ tensors $\m w$ that
%     generate the lowest-lying (unnormalized) $r$-magnon states
%     $\ket{\m w,m} \equiv \O\p{\m w, \sigma_\z^{\otimes{r}}} \ket{m}$
%     with definite excitation energy $\Delta$ above the PS manifold
%     $\M_0$ in a periodic 1D lattice of $N=101$ spins interacting via
%     the SU(2)-symmetric Hamiltonian in \eqref{eq:H_SU2} with power-law
%     couplings $h_{ij}=-1/\oper{dist}\p{i,j}^2$.  Here
%     $\oper{dist}\p{i,j}\equiv\min\p{\abs{i-j},N-\abs{i-j}}$ is the
%     minimal distance between spins $i,j$.  Top row ($r=2$) shows the
%     full 2D matrix $\m w = \sum_{i,j}w_{ij}\op{i}{j}$, and bottom row
%     ($r=3$) shows a single 2D slice of the full 3D tensor,
%     $\tilde{\m w}_0 = \sum_{i,j}w_{ij,0}\op{i}{j}$.  Blue (red) pixels
%     correspond to matrix elements with a positive (negative) sign, and
%     darker colors indicate larger absolute values.  In the case of
%     $r=2$, translational invariance enforces $w_{ij}=w_{i-d,j-d}$ for
%     all displacements $d$, so every row of $\m w$ is some translation
%     of a single (arbitrary) row.  Similarly, when $r=3$ every 2D slice
%     of the full 3D tensor can be obtained by appropriately translating
%     the 2D slice $\tilde{\m w}_0$ shown above.  In the case of SU($n$)
%     with $n>2$, the ``$r$-magnon'' excitation energies $\Delta$ and
%     tensors $\m w$ remain identical, but now correspond to a
%     degenerate set of states that are generated by different choices
%     of $r$-body operators in place of $\sigma_\z^{\otimes r}$.}
%   \label{fig:structure}
% \end{figure}

% focus on energy -- how different from non-interacting magnons?
% plot dispersion relation
% lax formulation for single-particle terms

The label $m$ suffices to identify the PS (Dicke) manifold
$\M_0=\oper{span}\set{\ket{m}:m\in\A_2\p{N}}$, but we need additional
labels to uniquely identify states in the $r$-magnon manifold $\M_r$
with $r>0$.  Single-magnon ($r=1$) spin-wave states of SU(2) spin
systems can be written in the form
\begin{align}
  \ket{m,\m v,\sigma_z}
  \equiv \O\p{\m v,\sigma_\z} \ket{m}
  = \sum_{j=1}^N v_j \sigma_\z^{(j)} \ket{m},
\end{align}
where each $v_j$ is a scalar and $\sigma_\z$ is a Pauli-$z$ operator.
If an SU(2) spin system is translationally invariant, then these
spin-wave eigenstates have $v_j=e^{\i k_{\m v}\c j}$ with wavenumber
$k_{\m v}\ne0$.  Similarly, $r$-magnon states for SU(2) take the form
$\ket{m,\m w,\sigma_\z^{\otimes r}} \equiv \O\p{\m
  w,\sigma_\z^{\otimes r}}\ket{m}$, where now $\m w$ is an $r$-index
tensor that encodes the structure of correlations in the excited state
$\ket{m,\m w,\sigma_\z^{\otimes r}}$.  Considering states of this
form, built using tensor products of Pauli-$z$ ($\sigma_\z$)
operators, suffices to construct a basis for all excited states of an
SU(2) spin system (see Appendix \ref{sec:operator_content}).  Note
that $r$-body operators with arbitrary coefficients $\m w$ will
generally couple the PS manifold $\M_0$ to all $q$-magnon manifolds
$\M_q$ with $q\le r$ (see Figure \ref{fig:manifold_sketch}).  In order
to generalize the notion of $r$-magnon manifolds $\M_r$ to the case of
SU($n$) with $n>2$, where the analogous notion of a spin length $S$ is
less intuitive, we can define $\M_r$ for $r>0$ as the manifold of
states that (i) can be generated by the action of $r$-body operators
on $\M_0$, and (ii) are orthogonal to all states in
$\bigoplus_{q<r}\M_q$.

@@

By repeating the above process incrementally for all
$r\le r_{\t{max}}$ (starting with $r=1$), we can construct an
orthonormal basis
$\col\B_{r_{\t{max}}}\equiv\bigotimes_{r\le r_{\t{max}}}\B_r$ for the
combined manifold of all states with at most $r_{\t{max}}$ magnons,
$\col\M_{r_{\t{max}}}\equiv\bigotimes_{r\le r_{\t{max}}}\M_r$.  The
purpose of constructing the basis $\col\B_{r_{\t{max}}}$ is to
simulate the dynamics induced by the addition of a spin operator
$\O\p{\m v,Y}$ to the base Hamilonian $H_0$, such that the total
Hamiltonian is $H = H_0 + \O\p{\m v,Y}$.  Throughout the construction
of $\col\B_s$, an appropriate choice of tensors $\m w$ ensures that
all states $\ket{m,\m w,X}\in\col\B_s$ are eigenstates of the base
Hamiltonian $H_0$ (see Appendix \ref{sec:eigenstates}), which is to
say that $H_0$ is diagonal with respect to $\col\B_s$.  Matrix
elements of $\O\p{\m v,Y}$, meanwhile, take the form
\begin{align}
  \bk{\ell,\m u,X|\O\p{\m v,Y}|m,\m w,Z}
  = \bk{\ell|\O\p{\m u,X}\O\p{\m v,Y}\O\p{\m w,Z}|m}.
\end{align}
In order to compute these matrix elements, we simplify general
products of multi-body operators in Appendix
\ref{sec:operator_product}.  For brevity, we will refer to numerical
simulations restricted to $\col\M_{r_{\t{max}}}$ as {\it
  order-$r_{\t{max}}$ truncated shell} (TS$_{r_{\t{max}}}$)
simulations.

An $M$-spin operator $\O\p{\m v,Y}$ can only couple $r$-magnon states
to $q$-magnon states with $\abs{q-r}\le M$.  In the perturbative
regime, $\norm{\O\p{\m v,Y}}<\Delta_{\t{gap}}/2$, TS$_{r_{\t{max}}}$
simulations of initially PS (zero-magnon) states therefore capture all
virtual processes through order $2\floor{r_{\t{max}}/M}+1$ in
perturbation theory.  When $\norm{\O\p{\m v,Y}}\ge\Delta_{\t{gap}}/2$,
perturbation theory formally breaks down, and the error incurred by
neglecting all $s$-magnon states with $s>r_{\t{max}}$ is not clear a
priori.  Nonetheless, TS$_{r_{\t{max}}}$ simulations can be
benchmarked by examining the populations of all $r$-magnon manifolds
$\M_r$ with $r\le r_{\t{max}}$.  As long as the maximal populations
within large-$r$ manifolds $\M_r$ are small, one should expect these
simulations to faithfully capture (at least qualitatively) the
dynamical behavior of the spin system under consideration.

@@

While the above restrictions on $z$ eliminate some trivial and
redundant states from the basis $\set{\ket{m,\m w,z}}$ for $\M_r$,
they do not guarantee that there will be {\it no} trivial states or
redundancies left over.  Furthermore, even after restrictions the
remaining basis is not orthogonal, which is to say that generally
$\bk{m,\m v,y|m,\m w,z}\ne0$ even when $y\ne z$.  In order to
construct an orthonormal basis for $\M_r$, we therefore need to build
a matrix $B_m$ with matrix elements
\begin{align}
  B_m\p{\m v,y;\m w,z} \equiv \bk{m,\m v,y|m,\m w,z}
  = \bk{m| \O\p{\m v,X_y}^\dag \O\p{\m w,X_z} |m},
\end{align}
where $y,z$ satisfy appropriate restrictions, and $\m v,\m w$ are
chosen from the set of tensors that generate eigenstates of $H_0$.
Eigenvectors of $B_m$ are then mutually orthogonal vectors of the form
$\sum_{\m w,z}c_{\m w z}\ket{m,\m w,z}$, whose squared norms are equal
to their corresponding eigenvalues.  Note that the matrix $B_m$ has a
block-diagonal structure, as it cannot (for example) couple states
$\ket{m,\m v,y},\ket{m,\m w,z}$ whose tensors $\m v,\m w$ generate
states with different interaction energies.  We therefore only need to
build and diagonalize this matrix independently within each block.

@@

\red{[TODO: write introductory text for this section]}

>>>>

@@

In this work, we generalize some of the techniques used to harness
SU(2)-symmetric spin interactions for metrological advantage to the
case of SU($n$).  In doing so, we develop a low-energy effective
theory of gapped SU($n$) ferromagnets with weakly broken SU($n$)
symmetry (e.g.~due to external fields or additional interaction
terms).  The restriction of our theory to ferromagnetic interactions
can be relaxed to the assumption of
\begin{enumerate*}
\item an initial state in the manifold of permutationally symmetric
  states, a multilevel generalization of the Dicke manifold for
  two-level spins, together with the assumption that
\item this manifold is gapped away from states that break
  permutational symmetry.
\end{enumerate*}
In addition to working out the effective Hamiltonian induced by weak
SU($n$) symmetry breaking through second order in perturbation theory,
we develop a diagrammatic method to systematically compute matrix
elements of arbitrary spin operators within manifolds of weakly broken
permutational symmetry, thereby enabling the numerical simulation of
dynamics induced by non-perturbative modifications to the SU($n$)
Heisenberg model.  While these simulations generally accumulate
uncontrolled errors from the truncation of Hilbert space, we identify
criteria under which numerical results correctly capture the
qualitative dynamical behavior of a system.

@@

If the coefficients $h_{ij}$ of the interaction Hamiltonian $H_0$ are
all negative, then the ground-state manifold $\M_0$ of $H_0$ consists
of spin-polarized, {\it permutationally symmetric} (PS) states that
are simultaneous $+1$ eigenstates of all permutation operators
$\Pi_{ij}$.  We will assume that all $h_{ij}<0$ throughout these
notes, but note that this restriction can be relaxed to the assumption
that the initial state of all spins is a permutationally symmetric
state (e.g.~a spin-polarized state).  We will also assume that the PS
manifold $\M_0$ is gapped away from all orthogonal states by an
interaction energy $\Delta_{\t{gap}}>0$.  Power-law couplings of the
form $h_{ij}\sim-1/\abs{i-j}^\alpha$ on a $D$-dimensional lattice, for
example, always yield a non-vanishing spectral gap $\Delta_{\t{gap}}$
for long-range interactions with $\alpha\le D$ (see Appendix
\ref{sec:spectral_gap}).

@@

If the addition $\O\p{\m w,X}$ to the SU($n$)-symmetric Hamiltonian
$H_0$ in \eqref{eq:H_0} is sufficiently small, namely with operator
norm $\norm{\O\p{\m w,X}}$ less than half the spectral gap
$\Delta_{\t{gap}}$ of $H_0$, $\norm{\O\p{\m w,X}}<\Delta_{\t{gap}}/2$,
then we can treat the effect of $\O\p{\m w,X}$ on the ground-state PS
manifold $\M_0$ perturbatively.  The effective Hamiltonians
$H_{\t{eff}}^{(1)}$ and $H_{\t{eff}}^{(2)}$ induced by $\O\p{\m w,X}$
on $\M_0$ at first and second order in perturbation theory
are\cite{bravyi2011schrieffer}
\begin{align}
  H_{\t{eff}}^{(1)} = \P_0 \O\p{\m w,X} \P_0,
  &&
  H_{\t{eff}}^{(2)} = - \sum_{\Delta\ne0}
  \f1\Delta \P_0 \O\p{\m w,X} \P_\Delta \O\p{\m w,X} \P_0,
  \label{eq:H_eff}
\end{align}
where $\P_\Delta$ is a projector onto the eigenspace $\E_\Delta$ of
the interaction Hamiltonian $H_0$ with interaction energy $\Delta$
above that of PS manifold $\M_0$.  In particular, $\P_0$ is a
projector onto $\M_0$ itself, and $\E_0=\M_0$.  The first order
effective Hamiltonian $H_{\t{eff}}^{(1)}$ simply projects
$\O\p{\m w,X}$ onto the PS manifold $\M_0$, and takes the form
\begin{align}
  H_{\t{eff}}^{(1)} \EQPS \mean w \, \col X,
  \label{eq:H_eff_1}
\end{align}
where $\EQPS$ denotes equality under a restriction to $\M_0$; the
coefficient $\mean w$ is the average of all $w_k$; and $\col X$ is a
collective version of $X$:
\begin{align}
  \mean w \equiv \f1{\abs{\D_N\p{M}}}
  \sum_{k\in\D_N\p{M}} w_k,
  &&
  \col X \equiv \sum_{k\in\D_N\p{M}} X_k,
  \label{eq:mean_col}
\end{align}
with $\abs{\D_N\p{M}}=\prod_{j=0}^{M-1}\p{N-j}$.  In the case of a
magnetic field $s_\z$ or Ising interactions $s_\z\otimes s_\z$, for
example,
\begin{align}
  \col{s_\z} = \sum_i s_\z^{(i)} = S_\z,
  &&
  \col{s_\z\otimes s_\z}
  = \sum_{i\ne j} s_\z^{(i)} s_\z^{(j)}
  = S_\z^2 - \f14 N.
\end{align}
The second order effective Hamiltonian $H_{\t{eff}}^{(2)}$ in
\eqref{eq:H_eff} takes more work to simplify due to the presence of a
projector $\P_\Delta$ onto the manifold $\E_\Delta$ of states with
excitation energy $\Delta$.  This projector essentially picks off the
part of $\O\p{\m w,X}$ that is strictly off-diagonal with respect to
the ground- and excited-state manifolds $\E_0$ and $\E_\Delta$.  We
therefore need to decompose $\O\p{\m w,X}$ into components that
generate states of definite excitation energy when acting on PS states
$\ket\psi\in\M_0$.  The SU($n$) symmetry of $H_0$ enables such a
decomposition to take the form
\begin{align}
  H_0 \O\p{\m w,X} \ket\psi
  = \sum_\Delta \p{E_0+\Delta} \O\p{\m w_\Delta,X} \ket\psi,
  \label{eq:excitation}
\end{align}
where, thinking of the tensor $\m w$ as a $\abs{\D_N\p{M}}$-component
vector, the tensor $\m w_\Delta$ can be found by diagonalizing a
matrix $\hat{\m h}$ of dimensions
$\abs{\D_N\p{M}}\times\abs{\D_N\p{M}}\sim N^M\times N^M$, and
projecting $\m w$ onto the eigenspace of $\hat{\m h}$ with eigenvalue
$\Delta$ (see Appendix \ref{sec:eigenstates}).  The matrix
$\hat{\m h}$ is determined entirely by the couplings couplings
$h_{ij}$ of $H_0$.  While the computational cost of diagonalizing
$\hat{\m h}$ is nominally $O\p{N^{3M}}$, spatial symmetries of $H_0$
and $\O\p{\m w,X}$ such as translational invariance or isotropy can
drastically reduce the complexity of this eigenvalue problem, as we
discuss in Appendix \ref{sec:symmetries}.

@@

In order to find the effective Hamiltonian $H_{\t{eff}}^{(2)}$ induced
by the $M$-body operator $\O\p{\m w,X}$ on the PS manifold $\M_0$ at
second order in perturbation theory, we constructed states with
definite excitation energy $\Delta>0$ with respect to $H_0$.  Rather
than treating these excited states as ``intermediate'' or ``virtual''
for the derivation of an effective ground-state Hamiltonian, we can
perform numerical simulations in the joint Hilbert space of both
ground and excited states.  When the perturbation $\O\p{\m w,X}$ is
small, namely with operator norm
$\norm{\O\p{\m w,X}}<\Delta_{\t{gap}}/2$, such simulations are
equivalent (up to perturbative corrections) to a higher-order
perturbation theory.  Moreover, the first-class treatment of excited
states in finite systems enables simulating the dynamics induced by
{\it non-perturbative} additions $\O\p{\m w,X}$ to the base
Hamiltonian $H_0$.

The basic idea will be to identify the set of states (manifolds)
$\M_r$ with a definite number ($r$) of spin-wave excitations, or {\it
  magnons}.  The ground-state manifold of PS states is simply the
zero-magnon manifold $\M_0$.  If we construct a basis for all $\M_r$
with $r\le r_{\t{max}}$, we can perform numerical simulations in the
combined manifold of all states with at most $r_{\t{max}}$ magnons; we
refer to such simulations as {\it order-}$r_{\t{max}}$ {\it truncated
  shell} (TS$_{r_{\t{max}}}$) simulations.  An $M$-spin operator
$\O\p{\m w,X}$ can only couple $r$-magnon states (in $\M_r$) to
$q$-magnon states (in $\M_q$) with $\abs{q-r}\le M$.  For initially PS
states, TS$_{r_{\t{max}}}$ simulations thus capture all virtual
processes through order $2\floor{r_{\t{max}}/M}+1$ in perturbation
theory, when $\norm{\O\p{\m w,X}}<\Delta_{\t{gap}}/2$.  When
$\norm{\O\p{\m w,X}}\ge\Delta_{\t{gap}}/2$, perturbation theory
formally breaks down, and the error incurred by neglecting all
$s$-magnon states with $s>r_{\t{max}}$ is not clear a priori.
Nonetheless, TS$_{r_{\t{max}}}$ simulations can be benchmarked by
examining the populations of all $r$-magnon manifolds $\M_r$ with
$r\le r_{\t{max}}$.  As long as the maximal populations within
large-$r$ manifolds $\M_r$ are small, one should expect
TS$_{r_{\t{max}}}$ simulations to faithfully capture (at least
qualitatively) the dynamical behavior of the spin system under
consideration.

The remainder of this section is devoted to the construction of a
basis for the $r$-magnon manifolds $\M_r$, and to the expansion of
general spin operators $\O\p{\m w,X}$ in this basis.  We begin with a
classification of excited states in the case of SU(2).  A common
starting point is to first classify states by a definite
(non-negative) spin length $S=\set{N/2,N/2-1,\cdots}$, and a spin
projection $m_\z\in\set{S,S-1,\cdots,-S}$ onto a fixed quantization
axis.  In fact, the manifold of states with definite fixed spin length
$S$ is precisely the $\p{N/2-S}$-magnon manifold $\M_{N/2-S}$.  The PS
manifold of Dicke states with $S=N/2$ is then $\M_0$, and more
generally the manifold of $r$-magnon states with $S=N/2-r$ is $\M_r$.
Rather than the spin projection $m_\z$, we then identify states by the
occupation numbers $m=\p{m_\up,m_\dn}\in\A_2\p{N}$, where $m_\mu$
denotes the total number of individual spins in state $\mu$, with
$m_\up+m_\dn=N$ and $m_\up-m_\dn=m_\z$.

The label $m$ suffices to identify the PS (Dicke) manifold
$\M_0=\oper{span}\set{\ket{m}:m\in\A_2\p{N}}$, but we need additional
labels to uniquely identify states in the $r$-magnon manifold $\M_r$
with $r>0$.  Single-magnon ($r=1$) spin-wave states of SU(2) spin
systems can be written in the form
\begin{align}
  \ket{m,\m v,\sigma_z}
  \equiv \O\p{\m v,\sigma_\z} \ket{m}
  = \sum_{j=1}^N v_j \sigma_\z^{(j)} \ket{m},
\end{align}
where each $v_j$ is a scalar and $\sigma_\z$ is a Pauli-$z$ operator.
If an SU(2) spin system is translationally invariant, then spin-wave
eigenstates have $v_j=e^{\i k_{\m v}\c j}$ with wavenumber
$k_{\m v}\ne0$.  Similarly, $r$-magnon states for SU(2) take the form
$\ket{m,\m w,\sigma_\z^{\otimes r}} \equiv \O\p{\m
  w,\sigma_\z^{\otimes r}}\ket{m}$, where now $\m w$ is an $r$-index
tensor that encodes the structure of correlations in the excited state
$\ket{m,\m w,\sigma_\z^{\otimes r}}$.  The coefficient tensor $\m w$
determines the energy of the excited state
$\ket{m,\m w,\sigma_\z^{\otimes r}}$ through the expansion in
\eqref{eq:excitation}.  Considering states of this form, built using
tensor products of Pauli-$z$ ($\sigma_\z$) operators, suffices to
construct a basis for all excited states of an SU(2) spin system (see
Appendix \ref{sec:operator_content}).  Note that $r$-body operators
with arbitrary coefficients $\m w$ will generally couple the PS
manifold $\M_0$ to all $q$-magnon manifolds $\M_q$ with $q\le r$ (see
Figure \ref{fig:manifold_sketch}).  In order to generalize the notion
of $r$-magnon manifolds $\M_r$ to the case of SU($n$) with $n>2$,
where the analogous notion of a spin length $S$ is less intuitive, we
can define $\M_r$ for $r>0$ as the manifold of states that (i) can be
generated by the action of $r$-body operators on $\M_0$, and (ii) are
orthogonal to all states in $\bigoplus_{q<r}\M_q$.

@@

By repeating the above process incrementally for all
$r\le r_{\t{max}}$ (starting with $r=1$), we can construct an
orthonormal basis
$\col\B_{r_{\t{max}}}\equiv\bigotimes_{r\le r_{\t{max}}}\B_r$ for the
combined manifold of all states with at most $r_{\t{max}}$ magnons.
During the construction of $\col\B_s$, an appropriate choice of
tensors $\m w$ ensures that all states $\ket{m,\m w,X}\in\col\B_s$ are
eigenstates of the base Hamiltonian $H_0$ (see Appendix
\ref{sec:eigenstates}), which is to say that $H_0$ is diagonal with
respect to $\col\B_s$.  Matrix elements of general spin operators
$\O\p{\m v,Y}$, meanwhile, take the form
\begin{align}
  \bk{\ell,\m u,X|\O\p{\m v,Y}|m,\m w,Z}
  = \bk{\ell|\O\p{\m u,X}\O\p{\m v,Y}\O\p{\m w,Z}|m}.
  \label{eq:matrix_element}
\end{align}
In order to compute these matrix elements, we simplify general
products of multi-body operators in Appendix
\ref{sec:operator_product}.

% dispersion relation? magnon interactions?
% lax formulation for single-particle terms
% solve collective model? Casmir operators?

@@

While the above restrictions on $z$ eliminate some trivial and
redundant states from the basis $\set{\ket{m,\m w,z}}$ for $\M_r$,
they do not guarantee that there will be {\it no} trivial states or
redundancies left over.  Furthermore, even after restrictions the
remaining basis is not orthogonal, which is to say that generally
$\bk{m,\m v,y|m,\m w,z}\ne0$ even when $y\ne z$.  In order to
construct an orthonormal basis for $\M_r$, we therefore need to build
a matrix $B_m$ with matrix elements
\begin{align}
  B_m\p{\m v,y;\m w,z} \equiv \bk{m,\m v,y|m,\m w,z}
  = \bk{m| \O\p{\m v,X_y}^\dag \O\p{\m w,X_z} |m},
  \label{eq:basis_matrix}
\end{align}
where $y,z$ satisfy appropriate restrictions, and $\m v,\m w$ are
chosen from the set of tensors that generate eigenstates of $H_0$.
Eigenvectors of $B_m$ are then mutually orthogonal vectors of the form
$\sum_{\m w,z}c_{\m w z}\ket{m,\m w,z}$, whose squared norms are equal
to their corresponding eigenvalues.  Note that the matrix $B_m$ has a
block-diagonal structure, as it cannot (for example) couple states
$\ket{m,\m v,y},\ket{m,\m w,z}$ whose tensors $\m v,\m w$ generate
states with different interaction energies.  We therefore only need to
build and diagonalize this matrix independently within each block.

@@

The construction of a basis $\B_r$ for the $r$-magnon manifold $\M_r$
requires multiplying two multi-body operators
$\O\p{\m v,X_y}^\dag, \O\p{\m w,X_z}$, and projecting the result onto
the permutationally symmetric (PS) manifold $\M_0$, as in
\eqref{eq:basis_matrix}.  Similarly, computing matrix elements of
operators $\O\p{\m v,Y}$ with respect to the basis
$\col\B_{r_{\t{max}}}\equiv\bigotimes_{r\le r_{\t{max}}}\B_r$ requires
multiplying three multi-body operators and projecting the result onto
$\M_0$, as in \eqref{eq:matrix_element}.  In this section, we develop
a diagrammatic technique for the sorts of calculations, and work out
some special cases of common interest.
